# Practical Machine Learning: Project Writeup
========================================================

Datasets were provided using data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. The goal of this project is to predict the manner in which they did the exercise (as indicated by the "classe" field).

Data was downloaded from the following locations:

Training Dataset: https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv
Test Dataset: https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

The training dataset will be used for training and model validation. The test dataset will be used to demonstrate the correctness of the model for the project submission on Coursera.

Cite:

Velloso, E.; Bulling, A.; Gellersen, H.; Ugulino, W.; Fuks, H. Qualitative Activity Recognition of Weight Lifting Exercises. Proceedings of 4th International Conference in Cooperation with SIGCHI (Augmented Human '13) . Stuttgart, Germany: ACM SIGCHI, 2013.

http://groupware.les.inf.puc-rio.br/har#ixzz352L6ksUA

Load the data sets:

```{r, cache = TRUE}
trainingSet <- read.csv("pml-training.csv")
testingSet <- read.csv("pml-testing.csv")
```

Next step is to partition the training data into a training set and a validation set. I decided to use 60% of the data for training and 40% for validation.

```{r, cache = TRUE}
library(caret)

testIndex <- createDataPartition(trainingSet$classe, p = 0.60, list=FALSE)

trainingData <- trainingSet[testIndex,]
validationData <- trainingSet[-testIndex,]
```

Now, create a parallel training control with five resampling iterations for the next step.
```{r}
library(doParallel)

nodes <- detectCores()
cl <- makeCluster(nodes)
registerDoParallel(cl)

trainingControl <- trainControl(method='cv', number = 5, allowParallel = TRUE)
```

Here the model fit is created. The method used is "random forest", and only the data fields used in the provided testing data set were used to create the model.
```{r cache = TRUE}
#head(testingSet)

modelFit <- train(classe ~ roll_belt + pitch_belt + yaw_belt + total_accel_belt
                  + gyros_belt_x + gyros_belt_y + gyros_belt_z
                  + accel_belt_x +  + accel_belt_y + accel_belt_z
                  + magnet_belt_x + magnet_belt_y + magnet_belt_z
                  + roll_arm + pitch_arm + yaw_arm + total_accel_arm
                  + gyros_arm_x + gyros_arm_y + gyros_arm_z
                  + accel_arm_x + accel_arm_y + accel_arm_z
                  + magnet_arm_x + magnet_arm_y + magnet_arm_z
                  + roll_dumbbell + pitch_dumbbell + yaw_dumbbell
                  + gyros_dumbbell_x + gyros_dumbbell_y + gyros_dumbbell_z
                  + accel_dumbbell_x + accel_dumbbell_y + accel_dumbbell_z
                  + magnet_dumbbell_x + magnet_dumbbell_y + magnet_dumbbell_z
                  + roll_forearm + pitch_forearm + yaw_forearm
                  + gyros_forearm_x + gyros_forearm_y + gyros_forearm_z
                  + accel_forearm_x + accel_forearm_y + accel_forearm_z
                  + magnet_forearm_x + magnet_forearm_y + magnet_forearm_z,
                  method = "rf", data = trainingData, trControl = trainingControl)

stopCluster(cl)
```

Training accuracy: `r modelFit$results[2,2] * 100`%
Training errors: `r 100 - (modelFit$results[2,2] * 100)`%

Now, test against the validation set:
```{r cache = TRUE}
validationPredict <- predict(modelFit, validationData)
validationAccuracy <- postResample(validationPredict, validationData$classe)
```

The actual out of sample error is `r (1-as.numeric(validationAccuracy[1])) * 100`%
